Interpreting Random Forest Predictions for Firearm Identification Using LIME

### 2019-01-24 paper/Interpreting Random Forest Predictions for Firearm Identification Using LIME.Rmd
worked on adding rfscore tree based bins

```{r setup, include = FALSE}
# Set knitr options
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "center")

# Load libraries
library(tidyverse)
library(bulletr)
library(randomForest)
library(lime)
library(cowplot)
library(purrr)
library(future)
library(furrr)

@@

```{r}
# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(rtrees$importance)
```

@@

  # Specify the options to use with lime
  hamby224_lime_inputs <- list(bin_continuous = c(rep(TRUE, 15), rep(FALSE, 2)),
                               quantile_bins = c(rep(TRUE, 5), rep(FALSE, 5), 
                                                 rep(TRUE, 5), rep(TRUE, 2)),
                               nbins = c(rep(2:6, 3), rep(4, 2)),
                               use_density = c(rep(TRUE, 15), TRUE, FALSE),
                               bin_method = c(rep("quantile_bins", 5),
                                              rep("equally_spaced", 5),
                                              rep("tree", 5),
                                              rep("quantile_bins", 2)))

@@

  # Apply lime to the full training data with the specified input options
  hamby224_lime_explain <- future_pmap(.l = hamby224_lime_inputs,
             .f = run_lime, # run_lime is one of my helper functions
             train = hamby173and252_train %>% select(rf_features),
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             response = hamby173and252_train$samesource,
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3,
             seed = TRUE)

@@

  # Name the items in the lime list
  names(hamby224_lime) <- map_chr(1:length(hamby224_lime_inputs), function(case) 
      sprintf("case: bin_continuous = %s, quantile_bins = %s, nbins = %0.f, use_density = %s, bin_method = %s",
              hamby224_lime_inputs$bin_continuous[case],
              hamby224_lime_inputs$quantile_bins[case],
              hamby224_lime_inputs$nbins[case],
              hamby224_lime_inputs$use_density[case],
              hamby224_lime_inputs$bin_method[case]))

  hamby224_lime_inputs <- hamby224_lime_inputs %>%
    unlist() %>%
    matrix(ncol = length(hamby224_lime_inputs), 
           dimnames = list(NULL, names(hamby224_lime_inputs))) %>%
    as.data.frame() %>%
    mutate(case = 1:length(hamby224_lime_inputs$quantile_bins)) %>%
    select(case, bin_continuous:bin_method)

@@

  # Join the data and the explanations and edit and add additional variables
  # Create the feature bin labels using my function "bin_labeller"
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature_bin = pmap_chr(list(feature = feature, 
                                  feature_value = feature_value,
                                  b_c = bin_continuous,
                                  q_b = quantile_bins,
                                  n_b = nbins,
                                  u_d = use_density,
                                  b_m = bin_method),
                            .f = bin_labeller, # bin_labeller is one of my helper functions
                            bin_data = hamby224_bin_boundaries,
                            case_info = hamby224_lime_inputs)) %>%
    mutate(feature = factor(feature),
           nbins = factor(nbins),
           feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE) %>%
    arrange(nbins)

@@

  # Reorder the variables of feature_desc and feature_bin for plotting purposes
  hamby224_test_explain <- hamby224_test_explain %>%
    mutate(feature_desc = reorder(feature_desc, strictly_less),
           feature_desc = reorder(feature_desc, feature_number),
           feature_desc = reorder(feature_desc, as.numeric(feature))) %>%
    select(bin_continuous:bin_method, case:feature_desc, feature_bin:strictly_less, 
           data, prediction)

@@

  # Create a data frame with the interesting information relating to the different
  # evaluations of lime and compute the difference and mean between the rf and rr
  # model predictions
  hamby224_lime_comparisons <- hamby224_test_explain %>%
    select(-data, -prediction) %>%
    group_by(case, bin_continuous, quantile_bins, nbins, use_density, bin_method) %>%
    slice(1) %>%
    ungroup() %>%
    select(bin_continuous, quantile_bins, nbins, use_density, 
           set, case, rf_features, rfscore, model_prediction, model_r2) %>%
    mutate(diff = rfscore - model_prediction,
           mean = (rfscore + model_prediction) / 2)

>>>>

```{r setup, include = FALSE}
# Set knitr options
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, 
                      fig.pos = "center")

# Load libraries
library(bulletr)
library(cowplot)
library(furrr)
library(future)
library(lime)
library(purrr)
library(randomForest)
library(tidyverse)
library(tree)

@@

```{r}
# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(rtrees$importance)

# Add the random forest score to the training set
hamby173and252_train$rfscore = predict(rtrees, 
                                       hamby173and252_train %>% select(rf_features), 
                                       type = "prob")[,2]
```

@@

  # Specify the options to use with lime
  hamby224_lime_inputs <- list(bin_continuous = c(rep(TRUE, 20), rep(FALSE, 2)),
                               quantile_bins = c(rep(TRUE, 5), rep(FALSE, 5), 
                                                 rep(TRUE, 10), rep(TRUE, 2)),
                               nbins = c(rep(2:6, 4), rep(4, 2)),
                               use_density = c(rep(TRUE, 20), TRUE, FALSE),
                               bin_method = c(rep("quantile_bins", 5),
                                              rep("equally_spaced", 5),
                                              rep("tree", 10),
                                              rep("quantile_bins", 2)),
                               response = c(rep(NA, 10), rep("samesource", 5), 
                                            rep("rfscore", 5), rep(NA, 2)))

@@

  # Apply lime to the full training data with the specified input options
  hamby224_lime_explain <- future_pmap(.l = hamby224_lime_inputs,
             .f = run_lime, # run_lime is one of my helper functions
             features = rf_features,
             train = hamby173and252_train,
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3,
             seed = TRUE)

@@

  # Name the items in the lime list
  names(hamby224_lime) <- map_chr(1:length(hamby224_lime_inputs), function(case) 
      sprintf("case: bin_continuous = %s, quantile_bins = %s, nbins = %0.f, use_density = %s, bin_method = %s, response = %s",
              hamby224_lime_inputs$bin_continuous[case],
              hamby224_lime_inputs$quantile_bins[case],
              hamby224_lime_inputs$nbins[case],
              hamby224_lime_inputs$use_density[case],
              hamby224_lime_inputs$bin_method[case],
              hamby224_lime_inputs$response[case]))

  hamby224_lime_inputs <- hamby224_lime_inputs %>%
    unlist() %>%
    matrix(ncol = length(hamby224_lime_inputs), 
           dimnames = list(NULL, names(hamby224_lime_inputs))) %>%
    as.data.frame() %>%
    mutate(case = 1:length(hamby224_lime_inputs$quantile_bins)) %>%
    select(case, bin_continuous:response)

@@

  # Join the data and the explanations and edit and add additional variables
  # Create the feature bin labels using my function "bin_labeller"
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature_bin = pmap_chr(list(feature = feature, 
                                  feature_value = feature_value,
                                  b_c = bin_continuous,
                                  q_b = quantile_bins,
                                  n_b = nbins,
                                  u_d = use_density,
                                  b_m = bin_method,
                                  r_v = response),
                            .f = bin_labeller, # bin_labeller is one of my helper functions
                            bin_data = hamby224_bin_boundaries,
                            case_info = hamby224_lime_inputs)) %>%
    mutate(feature = factor(feature),
           nbins = factor(nbins),
           feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE) %>%
    arrange(nbins)

@@

  # Reorder the variables of feature_desc and feature_bin for plotting purposes and
  # create new variables of situation and bin_situation
  hamby224_test_explain <- hamby224_test_explain %>%
    mutate(feature_desc = reorder(feature_desc, strictly_less),
           feature_desc = reorder(feature_desc, feature_number),
           feature_desc = reorder(feature_desc, as.numeric(feature))) %>%
    mutate(nbins = as.numeric(as.character(nbins)),
           situation = ifelse(bin_continuous == TRUE & bin_method == "quantile_bins", 
                              sprintf("%.0f quantile bins", nbins),
                              ifelse(bin_continuous == TRUE & bin_method == "equally_spaced",
                                     sprintf("%.0f equally spaced bins", nbins),
                                     ifelse(bin_continuous == TRUE & bin_method == "tree" &
                                              response == "samesource",
                                            sprintf("%.0f samesource tree based bins", nbins),
                                            ifelse(bin_continuous == TRUE & bin_method == "tree" &
                                              response == "rfscore",
                                              sprintf("%.0f rfscore tree based bins", nbins),
                                              ifelse(bin_continuous == FALSE & 
                                                       use_density == TRUE, 
                                                     "kernel density", "normal approximation"))))) %>%
             fct_relevel("2 quantile bins", "3 quantile bins", "4 quantile bins",
                         "5 quantile bins", "6 quantile bins", "2 equally spaced bins",
                         "3 equally spaced bins", "4 equally spaced bins",
                         "5 equally spaced bins", "6 equally spaced bins",
                         "2 samesource tree based bins", "3 samesource tree based bins",
                         "4 samesource tree based bins", "5 samesource tree based bins",
                         "6 samesource tree based bins")) %>%
    mutate(bin_situation = ifelse(bin_method == "quantile_bins" & bin_continuous == TRUE,
                                "quantile bins",
                                ifelse(bin_method == "equally_spaced" & bin_continuous == TRUE,
                                       "equally spaced bins", 
                                       ifelse(bin_method == "tree" & bin_continuous == TRUE & 
                                                response == "samesource",
                                              "samesource tree based bins", 
                                              ifelse(bin_method == "tree" & bin_continuous == TRUE & 
                                                response == "rfscore",
                                                "rfscore tree based bins", "other"))))) %>%
    mutate(bin_situation = factor(bin_situation)) %>%
    select(situation, bin_situation, bin_continuous:response, case:feature_desc,
           feature_bin:strictly_less, data, prediction)

@@

  # Create a data frame with the interesting information relating to the different
  # evaluations of lime and compute the difference and mean between the rf and rr
  # model predictions
  hamby224_lime_comparisons <- hamby224_test_explain %>%
    select(-data, -prediction) %>%
    group_by(case, bin_continuous, quantile_bins, nbins, use_density, bin_method, response) %>%
    slice(1) %>%
    ungroup() %>%
    select(situation, bin_situation, bin_method, bin_continuous, quantile_bins, response, 
           nbins, use_density, set, case, rf_features, rfscore, model_prediction, model_r2) %>%
    mutate(diff = rfscore - model_prediction,
           mean = (rfscore + model_prediction) / 2)
